### 2016-02-09 Notes/response.txt
new images, working on the draft and response

Major comments:

- In my opinion the authors are over-discussing well known
degeneracies between cosmological parameters adding no new
information. They formulate the key question in the second-to-last
paragraph of the introduction but there is no answer at the end of the
reading. While acknowledging the importance of having priors on some
parameters to better constrain neutrinos and dark energy, I think the
authors should explain or highlight how those priors will be obtained,
i.e., which external experiments are needed and how they will help CMB
Stage-IV. Moreover, additional datasets like galaxies surveys have
already been proved to be useful by a number of papers in the
literature, the authors should highlight what is new in their results.
Following on these concerns, it is not clear to me if the authors have
an idea on how to get those improvements, for example how to get a
better prior on n_s with non-CMB data. I find the discussion simply
speculative. This impacts in particular the abstract, the results and
the conclusion sections.

The key issue here seems to be: How best to address the question: What external information can be used to improve projected constraints by breaking degeneracies? We think the referee is saying there are 2 parts to this answer: (i) quantify how much the added information on other parameters breaks degeneracies and (ii) give a sense of how we could get that information. The referee points out that we did not do a good job explaining the possibilities for (ii) and that at least qualitatively other papers have discussed degeneracies.

- The baseline dataset in all the results is using CMB+BAO which means
already CMB + external information. This is confusing, we are not
improving on CMB then but rather on a combination of CMB and some
galaxy data.

The referee is right; we were not clear that CMB+BAO includes current BAO constraints, from e.g., BOSS.
It doesn't make sense to us to exclude things we currently know, so we think it important to include these constraints as our baseline. For example that will strongly bias the foracasted performance of CMB S4 when considered without future BAO experiments like DESI.

- I cannot understand Figures 1,2,3,4. If the far right is the
expected CMB S4+Planck-Pol+BAO15 then the lines starting point should
correspond to prior/sigma = 1. How is it possible than adding a prior
on a parameter 3 times bigger than the sigma from CMB
S4+Planck-Pol+BAO15 (e.g., on Ho) gives a better constrain? If I think
about it in the likelihood space I am multiplying a distribution with
another one having 3 times its width and I do not expect to have any
impact on the combined result. Also, I do not understand why we should
consider adding a prior which is bigger than the baseline sigma from
CMB S4+Planck-Pol+BAO. What is shown on the right axis?

We agree that this was unecessarily complicated so have changed all the figures to absolute priors on the parameters. The referee's question still remains: why does a 3% prior on H_0 say help if the data itself constrain it to 1%. We stand by our result that it does help. Imagine a 1-D likelihood with one experiment giving a 1% constraint and another [here called "external"] a similar 1% constraint. When combining the 2, the constraint will be tighter, with the new error given by the inverse square root of the sum of the two inverse variances: [1/\sigma_1^2 + 1/\sigma_2^2]^{-1/2}. In the simple example above the external prior is equal to the internal constraint, so the total constraint is reduced by \sqrt{2}. In the example given by the referee [external constraint 3 times looser], there will still be improvement, albeit only by a factor of \sqrt{1 + 1/9}, ~5%. In multi-dimensional space, things are not that simple, but the plots show how things improve. To come back to our first response, we believe this information is extremely important for future planning. While it is true that people have studied degeneracies before, their impact has not been presented in this way. The very fact that the plots are not familiar indicates that we are introducing a new way, a productive way, of thinking about planning. We have added a paragraph in the text that attempts to explain this point more clearly and -- as noted -- have changed the x-axis on all plots to make things less confusing.

- I think that the claim of dramatically improving dark energy is
wrong. That authors themselves show that in the best case scenario
with lots of external information we will improve on w no more than a
factor of 2. More generally I think the discussion on dark energy need
to say more explicitly that the CMB is not the best probe of DE and it
will tell us something only in combination with other data.

Good point. We agree and we have toned down the language especially in the abstract and added the sentence: "For this reason CMB provides limited information about dark energy on its own. However CMB plays an important role in dark energy studies when combined with low redshift probes."'

- The references are not adequate in many cases, see minor comments
for details.

see below

- I encourage the authors to review the text:
- - adjectives like “dramatically” and “extremely” are not useful in
giving an idea of an experiment constraining power;

good point; removed these.

- - in the abstract the authors mention signal to noise of the CMB
polarization signal and I think they mean data;

got it; thanks

- - in the abstract high redshift large scale structure is wrong, the
CMB is a high redshift signal lensed by low redshift large scale
structure though;

right; we meant "high redshift" compared to other lensing data, but the wording is clearly bad. We've fixed by removing "high redshift"

"'"
- - in the introduction rephrase the 4th sentence. New experiments are
built to have a better sensitivity, it is not a success, it is meant
to happen;

not sure we understand this point

- - in the introduction rephrase the dark energy section following my
comment above.

done; as mentioned above

- Why do the authors use fsky=0.75? CMB S4 is expected to observe
roughly half the sky which, after galaxy removal, should correspond to
fsky ~ 0.4 .

Right; this is aggressive and the true value will be determined, but we chose this to be consistent with Ref. 11. Added a footnote: "The exact specifications of CMB-S4 are to be determined; we choose this value of fsky to afford easy comparison with the results of [11]."

- Keeping the helium fraction fixed is wrong, Yp varies with baryon
density AND Neff.

That's indeed what we did, but the old version of the text did not explain. We changed to:
"Similarly, when we vary $N_{\rm eff}$ we adjust the helium abundance to its canonical value from Big Bang Nucleosynthesis, which depends also on the baryon density."'

Minor comments:

- In the introduction: Ref. 4 is WMAP 1year results, in this context
the general WMAP experiment paper Bennett et al. 2003 or the final
mission results Bennet et al. 2012 would be more appropriate. Same
thing for Planck, cite the Blue Book and/or the 2015 Overview papers.
Ref. 5 is Maxima and Ref. 6 is Boomerang, they are not satellites and
are not mentioned in the text.

Thanks, Fixed with Bennett et al. 2003 and Blue book

- ACT and SPT did not follow DASI, they have measured temperature with
ACTPol and SPTPol going for polarization. The first two should go into
the above temperature discussion with references to latest results or
general papers describing the experiments, and the last two should
contribute to the polarization discussion.

We organized the discussion into ground base and satellite experiments. For this reason
we mentioned ACT and SPT toghether with DASI. WE added references to general papers
describing actpol and sptpol.

- In the introduction the cosmological dependence on neutrinos
properties jumps from number to masses to number again, it is
confusing please rephrase.
-------------

- Choose between Universe or universe.

Done (Universe)

- Define the cosmological parameters listed in the introduction.
-------------

- In the introduction “model LCDM” -> “LCDM model”

Done

- Section II, “additional parameter is required” -> “additional
parameter is required?”

Done

- Section II, the natural first “baby” step is not a step. The massive
neutrinos LCDM is the authors baseline model.

Good point, we rephrased it.

- Define FWHM.

Done.

- Section IIIA, cite DESI.

Done (http://arxiv.org/abs/1308.0847)

- Section IIIA, what do the authors mean with the “current”
free-streaming scale?

We removed the word current

- Section IIIA “power power spectrum is suppressed” -> “power spectrum
is suppressed”

Thanks, done.

- Section IIIA, what is the meaning of “since the active neutrino
number density are known in the standard model”?

We meant that what is probed by the CMB is the energy density of neutrinos which can be
turned into a total neutrino mass (in the massive case) by knowing the number density
E= m n
We rephrease the sentence cause the presence of the term standard model can cause confusion.

- Section IIIA, “a constraint on the the” -> “a constraint on the”

Done.

- Section IIIA, add a reference to oscillation experiments result.

Done.

- Section IIIA, measuring ell<50 is not a design goal but rather an
analysis one given that the problem is the atmospheric emission.

We rephrase it --> If CMB S4 will not be designed to measure scales l < 50 to avoid difficulties in dealing with atmospheric noise

- Section IIIA, with satellites need to cite LiteBird and Core/Prism,
with the balloons Ebex and then Class for large modes from the ground.

Done PRISM(http://arxiv.org/abs/1306.2259) LiteBIRD(http://arxiv.org/abs/1311.2847)
EBEX (http://arxiv.org/abs/1007.3672) CLASS (http://arxiv.org/abs/1408.4788)

- Fig. 5 caption, there is no red line in the plot
We fixed it.

- Last part of Section IV, cite LSST.

Done (http://arxiv.org/abs/1211.0310)

- Last part of Section IV, what do the authors mean with surveys
beyoond DESI and LSST? LSST is beyond DESI and I am not aware of
post-LSST surveys. Also, galaxy surveys will not help in constraining
n_s.
---------------------